**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**

### 2025-07-02 paper/manuscript.qmd
update

---
title: "**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**"
shorttitle: "Mint Validation"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
  - name: Ana Neves  
    orcid: 0009-0006-0020-7599
    role:
      - Investigation
      - Methodology
      - Project administration
      - Supervision
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
  - name: Giulia Poreiro
    orcid: 0000-0002-2343-5109
    role:
      - Investigation
      - Methodology
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
      :::
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    authorship-agreements: ~
abstract: |
  TO DO.
#   **Significance Statement**. TODO.
keywords: "Interoception questionnaire, interoceptive accuracy scale, MAIA, Mint Validation, Body Awareness"
numbered-lines: false
floatsintext: true
mask: false
bibliography: bibliography.bib
format:
  apaquarto-pdf:
    documentmode: jou
    keep-tex: true
  apaquarto-docx: default
---

@@

TODO: write general intro.

@@

Goal of study 1: to generate a lot of items, analyze its structure and reduce them to a balanced set of items.

@@

We recruited 760 English-speaking participants using Prolific\textcopyright. We excluded 191 for failing at least one attention check, and 10 based on measures significantly related to the probability of failing attention checks [namely, the multivariate distance obtained with the OPTICS algorithm, @theriault2024check]. The final sample includes 559 participants (age = 37.0 $\pm$ 12.2 [18, 77]; 50.8\% women; Country of residence: 63.86% UK, 26.65% USA).
This study was approved by the University of Sussex' Ethics Committee (ER/MB2021/1).

@@

Based on the two goals outlined for this scale, namely to include different interoceptive modalities, and to explicitly state the context of the interoceptive experience (e.g., whether negative or positive), we generated items in a systematic way following a combinatory approach, where each item's category corresponds to the combination of a specific modality and context (@fig-one).

```{r}
#| label: fig-one
#| fig-cap: "The conceptual grid used to generate the 120 initial items (top-left). Each item belong both to an interoceptive modalitiy and a facet, with the number of each item per category indicated in the circles. The askterisk denotes the additional presence of an attention check item in that category. In the experiment, these items were presented on different pages grouped either by modality (bottom-left), by facet (top-right), or entirely randomly. The Correlation Similarity (bottom-right) analysis suggested that the correlation matrix obtained from the participants assigned to the random-grouping condition was slightly more similar (but non-significantly) to the one obtained in the modality-grouping condition, suggesting that 1) the scale's structure is robust to different presentation conditions; 2) modality-grouping might tend to facilitate the emergence of the underlying item structure (and thus be interpreted as being more natural)."
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

@@

We identified 7 "modalities" (cardiac, respiratory, gastric, genital, skin & temperature, bladder & colon, and a "general state" category corresponding to a holistic and general awareness of an interoceptive state or dimension). Through iterative refinement (e.g., splitting or merging different categories together), we settled on 6 "facets", which encompass both *contexts* of experience (negative and positive arousal, namely anxious and sexual states), and potential distinct *mechanisms* (nociception & pleasure, sensitivity, accuracy, and confusion).

@@

To avoid presenting all the 120 items on a single long and discouraging page, we split them into different pages.
Participants were randomly assigned to one of three conditions, driving how items were grouped on the same page: 1) items grouped by modality (i.e., all cardiac items on the first page, all colon & bladder  items on the second, etc.), 2) items grouped by facet, or 3) items presented fully randomly (but balanced randomly across 6 pages). The order of the item on any given page and the order of the modalities/facets were randomized.
Each participant completed the full set of 120 items, with the 8 attention check items interspersed throughout.
The online experiment was implemented using JsPsych [@de2015jspsych], and item responses were recorded using 7-points Likert scales (0 = Disagree, 6 = Agree).

@@

Items deemed "redundant" (which can distort the item structure estimation by introducing multicollinearity or local dependencies) will be identified (using the recommended threshold of 0.25) using Unique Variable Analysis [UVA, @christensen2023unique], a novel and principled method derived from network psychometrics.

The structure of the items will be analyzed using the recently-developed Exploratory Graph Analysis [EGA, @golino2017exploratory] framework, which allows to jointly estimate the number of dimensions (i.e., clusters of items), the structure, as well as its stability using bootstrapping [@golino2020investigating]. At a fundamental level, EGA conceptualizes variables as nodes in a network, with connections (edges) reflecting associations between them. Evidence has underlined its suitability as an alternative to traditional factor analysis, addressing some of its limitations such as the assumption of a "latent" source of variability, issues in estimation of the optimal factor numbers, and poor performance in complex population structures, while remaining comparable and interpretable [@christensen2021equivalency; @jimenez2023dimensionality]. In particular, nodes communities (i.e., clusters of items) can be in practice interpreted as distinct "dimensions", similarly to traditional latent factors - but without explicitly assuming their existence [@christensen2021equivalency].

After removing redundant items using UVA, we will iteratively fit hierarchical EGA models (which additionally estimates higher-order "meta" clusters) using "glasso" [**REF**] and the "leiden" algorithm [**REF**] for community detection, refining the item pool at each step. We will start by removing items with a low (< 80%) cluster stability (i.e., volatile items which jump between clusters across bootstrapped samples), followed by odd items belonging to no clusters or pairs of items (i.e., we keep items belonging to clusters of more than 2 items). Finally, for each lower-level cluster, we will select the 3 items with the highest node centrality (i.e., the highest loading in the cluster).

@@

The correlation matrix similarity analysis yielded no significant differences between the similarity of the random-grouping condition with the modality-grouping and facet-grouping conditions ($PSI_{\text{Random vs. Facet}} = 0.81$, $PSI_{\text{Random vs. Modality}} = 0.82$, $p = .45$; $RVadj_{\text{Random vs. Facet}} = 0.77$, $RVadj_{\text{Random vs. Modality}} = 0.78$, $p = .74$; $SMI_{\text{Random vs. Facet}} = 0.49$, $SMI_{\text{Random vs. Modality}} = 0.51$, $p = .52$).

From the 120 initial items, UVA flagged 4 redundant items that we removed. We then removed 40 items that showed low cluster stability, and 9 items that were part of clusters with less than 3 items. Finally, We kept the 3 items with the highest loading in their lower-level structure (removing 13 items in the process), resulting in 54 items in the final item pool.

@@

Possible (although not at all significant) bias consistent across indices in favour of a greater similarity with the modality-grouping condition.

We grouped some of the dimensions under "Awareness", despite being similar to items included in the IAS typically conceptualized as "Accuracy". However, we believe that subjective scales cannot be pure measures of accuracy, as they are inherently subjective and thus reflect a degree of awareness (of one's accurate noticing or sensation).

@@

TODO: write intro.
Goal of study 2: to validate the Mint scale against other interoception scales.

@@

We recruited 921 English-speaking participants via Prolific\textcopyright and SONA, from which 118 were excluded for failing at least one attention check and 6 based on multivariate distance (using the same procedure as for study 1). 
60 participants were further excluded due to missing data following a technical error.
The final sample includes 737 participants (age = 36.8 $\pm$ 14.7 [18, 87]; 57.3\% women; Country of residence: 75.17% United Kingdom, 24.83% other).
This study was approved by the University of Sussex' Ethics Committee (ER/EB672/2).

@@

The experiment included the following questionnaires.

**TODO: What about attention checks?**

##### The Multimodal Interoception Questionnaire (Mint)

The 45 items of the Mint scale were presented in a random order, with the same 7-point Likert scale as in study 1 (0 = Disagree, 6 = Agree).

##### Multidimensional Assessment of Interoceptive Awareness - Second Version (MAIA-2)

The 37 items of the MAIA-2 [@mehling2018multidimensional] on a 7-point Likert scale from *Never* (0) to *Always* (6). The MAIA-2 consists of 8 scales: Noticing, Not-Distracting, Not-Worrying, Attention Regulation, Emotional Awareness, Self-Regulation, Body Listening and Trust. The scales address 5 dimensions of body awareness: Awareness of Body Sensations, Emotional Reaction and Attentional Response to Sensations, Capacity to Regulate Attention, Awareness of Mind-Body Integration, Trusting Body Sensations.

##### The Body Perception Questionnaire - Very Short Form (BPQ- VSF)

The 12 item BPQ-VSF [@cabrera2018assessing] related to bodily awareness on a 5-point Likert scale from *Never* (0) to *Always* (5).

##### The Interoceptive Accuracy Scale (IAS)

The 21 items of the IAS [@murphy2020testing] assessing interoceptive accuracy to physical sensations on a 5-point Likert scale ranging from *Disagree Strongly* (1) to *Strongly Agree*.

##### Toronto Alexithymia Scale (TAS-20)

The 20-items TAS [@leising2009toronto] assessing Alexithymia traits grouped into three dimensions: Difficulty Identifying Feelings, Difficulty Describing Feelings, and Externally Oriented Thinking. Items are rated in a 5-point Likert scale from *Strongly Disagree* (1) to *Strongly Agree*(5).

##### Emotion Reactivity Scale - Brief Version (B-ERS)

The 6-item B-ERS [@veilleux2024development] assessing emotional sensitivity on a 5-point Likert scale from *Not like me at all* (0) to *Extremely like me* (4).

##### Cognitive Emotion Regulation Questionnaire-short (CERQ-short)

The 18 items CERQ-short [@garnefski2006cognitive] on a 5-point Likert scale ranging from *Almost Never* (1) to *Almost Always* (5). This questionnaire measures eight cognitive responses to threatening or stressful events, including: Self-blame, Other-blame, Rumination, Catastrophizing, Putting into Perspective, Positive Refocusing, Positive Reappraisal, Acceptance, and Planning.

##### Černis Felt Sense of Anomaly scale - short-form (CEFSA-14)

The 14 item CEFSA [@vcernis2024measuring] assessing a specific sub-type of dissociative experiences in adolescence and adulthood - i.e., felt sense of anomaly-type
dissociation. the items are rated ON A 5-point Likert scale from *Never* (0) to *Always* (4), forming seven factors: Anomalous Experience of the Self, Anomalous Experience of the Body , Anomalous Experiences of Emotion , Altered Sense of Familiarity , Altered Sense of Connection, Altered Sense of Agency , and Altered Sense of Reality

##### Primal World Beliefs (PI-18)

the 18 items PI [@clifton2021brief] measuring highest-order primal world beliefs - i.e., beliefs about the basic character of the world: Good, Safe, Enticing, Alive. Responses are scored on a 6-point Likert scale from *Strongly Disagree* (0) to *Strongly Agree* (5).

##### Life Satisfaction, Depression and Anxiety (PHQ-4)

The 4 item scale of the PHQ [@kroenke2009ultra] assessing depression and anxiety, with responses scored as *Not at all* (0), *Once or twice* (0.5), *Several days* (1), *More than half of the days* (2), *nearly every day* (3). This revised version added as a response option (i.e., Once or twice) as a way to improve the scale's sensitivity to milder mood fluctuations [@makowskimeasuring].

<!-- PHQ-4 + LS: presented together -->

##### Mental Health

##### Somatic Health

##### Patient Health Questionnaire - Somatic Symptoms (PHQ-15)

The PHQ-15 [@kroenke2002phq] is comprised of 15 somatic symptoms, with responses scored as *Not at all* (0), *Bothered a little* (1) and *bothered a lot* (2).

Participants were asked to report on their use of wearable devices to monitor several health indices: heart rate, number of steps, calories burned, sleep quality, and weight (including via a regular scale). For each selected index, a question appeared asking them to rate how often they check that index; responses were scored as *Never*, *Very rarely*, *A few times per week*, *Once a day*, *A few times per day*, and *A few times per hour*. Additionally, participants were asked to rate how important this information was to them on an 8-point Likert scale ranging from *Not at all* (0) to *Very much* (7).

Questions regarding physical activity were also included. Participants were asked to rate how physically active they consider themselves on an 8-point Likert scale, from *Not at all* (0) to *Very much* (7), and to self-report how many hours of active workout they engage in per week.

#### Randomization

In order to avoid the repetition of similar types of questions and and balance longer and shorter questionnaires, we partitioned them into three groups (and randomized their order within them): (1) interoception questionnaires (MAIA, IAS, BPQ), (2) emotions (TAS, CERQ, ERS, PI-18), and (3) health (somatic symptoms, mental health, LS + PHQ-4, CEFSA). 
After completing demographic questions, participants always started with the Mint scale, and each following interoception questionnaire was interspersed with two questionnaires from the emotions and pathology groups.
In order to make the experiment more enjoyable, the experiment ended with a radar chart summarizing the participants' responses to the Mint scale^[The experiment can be tested by following the link on https://github.com/RealityBending/InteroceptionScale].

@@

We will start by confirming and further refining the structure of the Mint scale (see @fig-two) using the same EGA model as in Study 1. 
We will then compute the lower-level dimensions as well as the higher-level metaclusters' scores by averaging the items in each cluster.
The convergent validity of the final set of items will be assessed by computing the correlations between the Mint scale and the other interoception questionnaires (MAIA, BPQ, IAS).

```{r}
#| label: fig-two
#| fig-cap: "EGA stability and Hclust. TODO."
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

Next, we will test the predictive power of the Mint scale relative to the other interoception questionnaires.
We will first assess the importance of each interoceptive dimension (from all the scales) as a unique predictor by fitting regression models (linear for continuous measures - e.g., depression score - and logistic for binary variables - e.g., presence *vs.* absence of mood disorders) to predict various outcomes, and compare the best between the 4 interoception scales (based on the R2).
We will then assess the predictive performance of each scale as a whole by comparing regression models with all the dimensions entered as predictors (note that the IAS and BPQ only have one total score variable). We will evaluate models based on their total explained variance (R2), as well as on the Bayesian Information Criterion (BIC), which penalizes for predictor numbers, thus offering a balance between model parsimony and predictive power. For the logistic model, we will quantify the discriminative power by computing the Area Under the Curve (AUC) of the Receiver Operating Characteristic (ROC) curve, which assesses the model's discriminative power (the combination of sensitivity and specificity).

In order to evaluate the potential of a short version of our scale, we will compare 4 variants of the Mint models: the (full) Mint (including all lower-level dimensions), the "metaMint" (only including the metaclusters), the "miniMint" (including the metaclusters computed from a reduced number of items), and the "microMint" (including only the most representative dimension of each metacluster).
Moreover, we will also include an alternative "interoception-only" version of the MAIA ("iMAIA") that only includes the 3 most interoceptive dimensions (Body Listening, Noticing, and Emotional Awareness)^[See correlation results to further justify this selection.].

**TODO: Ana: do you think that this selection of dims for the iMAIA is fair? Or should I include other dimensions that you think are also very low-level and interoceptive?**

@@

The first application of EGA to the initial set of 42 items, which reproduced the expected 14 lower-level clusters of triplets and 3 higher-order metaclusters, with the exception of the *UrSe* (Urointestinal Sensitivity), for which one item moved to the *Olfa* (Olfactory Compensation) cluster.
In order to further balance and reduce the items, we removedt the *UrSe* cluster, as well as *StaS* (State Specificity; which in comparison to the other items stood out as containing vague and overlapping items (e.g., "Being sexually aroused is a very different bodily feeling compared to other states (e.g., feeling anxious, relaxed, or after physical exercise)"); *SexO* (Sexual Organs Sensitivity), which overlapped with the *SexS* (Sexual Arousal Sensitivity) cluster; and *CaNo* (Cardiorespiratory Noticing), which overlapped with the *CaCo* (Cardiorespiratory Confusion) cluster.

The final set of 36 items yielded a good fit (Generalized Total Entropy Fit Index = -77.23), with all items showing high cluster stability (> 90\%), with the exception of *Olfa*. The final structure included 11 lower-level clusters, grouped into 3 higher-order metaclusters: "Interoceptive Deficit" (containing *CaCo*, *UrIn*, *Derm*, *Sati*, *Olfa*), "Interoceptive Awareness" (containing *ExAc*, *RelA*, *SexS*), and "Visceroception" (containing *Card*, *Resp*, *Gast*). Item centrality (intepretable as loadings) are shown in @fig-three. Additionally, we also ran a hierarchical clustering analysis which replicated this structure (see @fig-two).

```{r}
#| label: fig-three
#| fig-cap: "Item loadings. TODO."
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

>>>>

---
title: "**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**"
shorttitle: "Mint Validation"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
  - name: Ana Neves  
    orcid: 0009-0006-0020-7599
    role:
      - Investigation
      - Methodology
      - Project administration
      - Supervision
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
  - name: Emma Benn 
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Maisie Bennett
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Giulia Poreiro
    orcid: 0000-0002-2343-5109
    role:
      - Investigation
      - Methodology
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
      :::
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    authorship-agreements: ~
abstract: |
  TO DO.
#   **Significance Statement**. TODO.
keywords: "Interoception questionnaire, interoceptive accuracy scale, MAIA, Mint Validation, Body Awareness"
numbered-lines: false
floatsintext: true
mask: false
bibliography: bibliography.bib
format:
  apaquarto-pdf:
    documentmode: jou
    keep-tex: true
  apaquarto-docx: default
---

@@

@@

- Goal of study 1: to generate a lot of items, analyze its structure and reduce them to a balanced set of items.

@@

We recruited 760 English-speaking participants using Prolific\textcopyright. We excluded 191 for failing at least one attention check, and 10 based on measures significantly related to the probability of failing attention checks [namely, the multivariate distance obtained with the OPTICS algorithm, @theriault2024check]. The final sample includes 559 participants (age = 37.0 $\pm$ 12.2 [18, 77]; 50.8\% women; country of residence: 63.86% UK, 26.65% USA).
This study was approved by the University of Sussex' Ethics Committee (ER/MB2021/1).

@@

Based on the two goals outlined for this scale, namely to include different interoceptive modalities, and to explicitly state the context of the interoceptive experience (e.g., whether negative or positive), we generated items in a systematic way following a combinatory approach, where each item's category corresponds to the union of a specific modality and context (@fig-one).

```{r}
#| label: fig-one
#| fig-cap: "The conceptual grid used to generate the 120 initial items (top-left). Each item belongs both to an interoceptive modality (vertical) and a facet (horizontal), with the number of each item per category indicated in the circles. The asterisk denotes the additional presence of an attention check item in that category. In the experiment, these items were presented on different pages grouped either by modality (bottom-left), by facet (top-right), or randomly. The Correlation Similarity (bottom-right) analysis suggested that the correlation matrix obtained from the participants assigned to the random-grouping condition was slightly more similar (but non-significantly) to the one obtained in the modality-grouping condition, suggesting that the scale's structure is robust to different presentation conditions, and that the modality-grouping might potentially tend to facilitate the emergence of the underlying item structure (and thus be interpreted as being more \"natural\")."
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

@@

We identified 7 "modalities" (cardiac, respiratory, gastric, genital, skin & temperature (thermoregulation), bladder & colon, and a "general state" category corresponding to a holistic and general awareness of an interoceptive state or dimension). Through iterative refinement (e.g., splitting or merging different categories together), we settled on 6 "facets", which encompass both *contexts* of experience (negative and positive arousal, e.g., anxious and sexual states), and potential distinct *mechanisms* (nociception & pleasure, sensitivity, accuracy, and confusion).

@@

To avoid presenting all the 120 items on a single long and discouraging page, we split them into different pages.
Participants were randomly assigned to one of three conditions, driving how items were grouped on the same page: 1) items grouped by modality (i.e., all cardiac items on the first page, all colon & bladder items on the second, etc.), 2) items grouped by facet, or 3) items presented fully randomly (but with their number balanced across 6 pages). The order of the item on any given page and the order of the modalities/facets was randomized.
Each participant completed the full set of 120 items, with 8 attention check items interspersed throughout.
The online experiment was implemented using JsPsych [@de2015jspsych], and item responses were recorded using 7-points Likert scales (0 = Disagree, 6 = Agree).

@@

Items deemed "redundant" (which can distort the item structure estimation by introducing multicollinearity or local dependencies) were identified (using the recommended threshold of 0.25) using Unique Variable Analysis [UVA, @christensen2023unique], a novel and principled method derived from network psychometrics.

The structure of the items was analyzed using the recently-developed Exploratory Graph Analysis [EGA, @golino2017exploratory] framework, which allows to jointly estimate the number of dimensions (i.e., clusters of items), the structure, as well as its stability using bootstrapping [@golino2020investigating]. At a fundamental level, EGA conceptualizes variables as nodes in a network, with connections (edges) reflecting associations between them. Evidence has underlined its suitability as an alternative to traditional factor analysis, addressing some of its limitations such as the assumption of a "latent" source of variability, issues in estimation of the optimal factor numbers, and poor performance in complex population structures, while remaining comparable and interpretable [@christensen2021equivalency; @jimenez2023dimensionality]. In particular, nodes communities (i.e., clusters of items) can be in practice interpreted as distinct "dimensions", similarly to traditional latent factors - but without explicitly assuming their existence [@christensen2021equivalency].

After removing redundant items using UVA, we iteratively fitted hierarchical EGA models (which additionally estimates higher-order "meta" clusters) using "glasso" [@friedman2008sparse] and the "Leiden" algorithm [@traag2019louvain] for community detection, refining the item pool at each step. We started by removing items with a low (< 80%) cluster stability (i.e., volatile items which jump between clusters across bootstrapped samples), followed by odd items belonging to no clusters or pairs of items (i.e., we keep items belonging to clusters of more than 2 items). Finally, for each lower-level cluster, we selected the 3 items with the highest node centrality (i.e., the highest loading in the cluster).

@@

The correlation matrix similarity analysis yielded no significant differences between the similarity of the random-grouping condition with the modality-grouping and facet-grouping conditions ($PSI_{\text{Random vs. Facet}} = 0.81$, $PSI_{\text{Random vs. Modality}} = 0.82$, $p = .45$; $RVadj_{\text{Random vs. Facet}} = 0.77$, $RVadj_{\text{Random vs. Modality}} = 0.78$, $p = .74$; $SMI_{\text{Random vs. Facet}} = 0.49$, $SMI_{\text{Random vs. Modality}} = 0.51$, $p = .52$), despite a consistent bias in favour of the modality-grouping condition.

From the 120 initial items, 4 redundant items were flagged by UVA. We then removed 40 items that showed low cluster stability, and 9 items that were part of clusters with less than 3 items. Finally, We kept the 3 items with the highest loading in their lower-level structure (removing 13 items in the process), resulting in 54 items in the final item pool.

@@

- Possible (although not at all significant) bias consistent across indices in favour of a greater similarity with the modality-grouping condition. Futher research with more data (here we only had ~200 observation per condition) might be able to yield more conclusive results. In general, testing the presentation mode is a novel approach which should be more widely applied to psychometric validations.
- We grouped some of the dimensions under "Awareness", despite being similar to items included in the IAS typically conceptualized as "Accuracy". However, we believe that subjective scales cannot be pure measures of accuracy, as they are inherently subjective and thus reflect a degree of awareness (of one's accurate noticing or sensation).
- Olfactory compensation: pseudo-interoceptive pathway whereby people with interoceptive deficits would compensate by gathering evidence about their internal state externally (e.g., via smell).
- The metacluster "Sensitivity" seems to be grouping iterms related to visceropcetion and (hyper)sensitivity (Derm).

@@

- Goal of study 2: to validate the Mint scale against other interoception scales.

@@

We recruited 921 English-speaking participants via SONA and Prolific\textcopyright, from which 118 were excluded for failing at least one attention check and 6 based on multivariate distance (using the same procedure as for study 1). 
60 participants were further excluded due to missing data following a technical issue.
The final sample includes 737 participants (age = 36.8 $\pm$ 14.7 [18, 87]; 57.3\% women; Country of residence: 75.17% United Kingdom, 24.83% other).
This study was approved by the University of Sussex' Ethics Committee (ER/EB672/2).

### Measures

```{r}
#| label: fig-two
#| fig-cap: "Structure analysis of the final set of 33 Mint items. Hierarchical Exploratory Graph Analysis (EGA) was applied to jointly identify clusters of items, as well as higher-order metaclusters. The reliability of each item can be quantified by the proportion of structure replication across bootstrapped samples, with high values indicative of high structural stability. The Mint scale displayed an excellent structural consistency, with the exception of the *Olfa* cluster which belonging to the *Deficit* metaclusters warrants further research. Bottom plot shows the result of hierarchical clustering, providing evidence for structure statbility across methods, and allowing for more granular understanding of the relationships between variables."
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "80%"

The 45 items of the Multimodal Interoception Questionnaire (Mint) scale were presented in a random order, with the same 7-point Likert scale as in study 1 (0 = *Disagree*, 6 = *Agree*).

The Multidimensional Assessment of Interoceptive Awareness [MAIA-2, @mehling2018multidimensional] measures 8 dimensions with 37 items presented on a 7-point Likert scale (0 = *Never*, 6 = *Always*). It includes bodily dimensions such as Body Listening, Noticing, Emotional Awareness, dimensions related to emotion regulation, such as Trust, Not-worrying, and Self-Regulation, as well as dimensions related to attention, such as Attention Regulation and Not-Distracting. The relationship of some of these dimensions to interoception remains debated, particularly Not-Distracting and Not-Worrying, which tend to show weak or non-existent correlations with the other MAIA subscales [@ferentzi2021examining].

The Body Perception Questionnaire - Very Short Form [BPQ-VSF, @cabrera2018assessing]  measures a general score of bodily awareness with 12 items presented on 5-point Likert scale (0 = *Never*, 5 = *Always*).

The Interoceptive Accuracy Scale [IAS, @murphy2020testing] measures a general score of interoceptive accuracy to physical sensations with 21 items presented on a 5-point Likert scale (1 = *Disagree Strongly*, 5 = *Strongly Agree*).

The Toronto Alexithymia Scale TAS [TAS-20, @leising2009toronto] measures 3 Alexithymia dimensions with 20 items 5-point Likert scale (1 = *Strongly Disagree*, 5 = *Strongly Agree*: Difficulty Identifying Feelings (DIF), Difficulty Describing Feelings (DDF), and Externally Oriented Thinking (EOR).

The Emotion Reactivity Scale - Brief Version [B-ERS, @veilleux2024development] measures 3 dimensions with 6 items on a 5-point Likert scale (0 = *Not like me at all*, 4 = *Extremely like me*): Arousal (*"My moods are very strong and powerful"*), Sensitivity (*"I tend to get very emotional very easily"*), and Persistence (*"When I am angry/upset, it takes me much longer than most people to calm down"*).

The Cognitive Emotion Regulation Questionnaire [CERQ-short, @garnefski2006cognitive] measures 9 adaptive and maladaptive emotion regulation strategies with 18 items (1 = *Almost Never*, 5 = *Almost Always*). The strategies include Self-blame, Other-blame, Rumination, Catastrophizing, Putting into Perspective, Positive Refocusing, Positive Reappraisal, Acceptance, and Planning.

The Černis Felt Sense of Anomaly scale - short form [CEFSA-14, @vcernis2024measuring] measures 7 dissociative experiences with 14 items on a 5-point Likert scale (0 = *Never*, 4 = *Always*). This scale is designed to measure dissociative experiences in adolescence and adulthood, specifically focusing on the felt sense of anomaly-type dissociation. It includes Anomalous Experience of the Self, Anomalous Experience of the Body, Anomalous Experiences of Emotion, Altered Sense of Familiarity, Altered Sense of Connection, Altered Sense of Agency, and Altered Sense of Reality.

The Primal World Beliefs Inventory [PI-18, @clifton2021brief] measures higher-order beliefs about the basic character of the world: Good, Safe, Enticing, and Alive. The 18 items are presented on a 6-point Likert scale (0 = *Strongly Disagree*, 5 = *Strongly Agree*).

The Single Item Life Satisfaction scale [SILS, @jovanovic2020longer] was followed by the PHQ-4 [@kroenke2009ultra] assessing depression and anxiety symptoms with 4 items. We used the refined version of the PHQ-4 [@makowski2025measuring], which adds an additional response option ("Once or twice"), increasing sensitivity to subclinical fluctuations.

Participants reported about owning any wearable devices to monitor health indices such as heart rate, number of steps, calories burned, sleep quality, and weight. For each selected device, a follow-up question inquired about the frequency of usage and subjective importance of that measure.

Participants were asked to rate how physically active they consider themselves and how many hours of active workout they engage in per week. Participant's BMI was computed using height and weight.

@@

In order to avoid the repetition of similar types of questions and and balance longer and shorter questionnaires, we partitioned the measures into three groups (and randomized the order within them): 1) interoception questionnaires (MAIA, IAS, BPQ), 2) emotions (TAS, CERQ, ERS, PI-18), and 3) health (somatic symptoms, mental health, LS + PHQ-4, CEFSA). 
After completing demographic questions, participants always started with the Mint scale, and each following interoception questionnaire was interspersed with two questionnaires from the emotions and pathology groups. 7 attention checks were embedded throughout (one within each major questionnaire).
In order to make the experiment more enjoyable, the experiment ended with a radar chart summarizing the participants' responses to the Mint scale^[The experiment can be tested by following the link on https://github.com/RealityBending/InteroceptionScale].

```{r}
#| label: fig-three
#| fig-cap: "Item loadings. The table shows each item of the Mint with its cluster centrality, which is the EGA equivalent (in terms of interpretation) of factor loadings. It also shows how each lower-level cluster is related to higher-order metaclusters."
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

@@

We started by confirming and further refining the structure of the Mint scale (see @fig-two) using the same EGA model as in Study 1. 
We then computed the lower-level dimensions and the higher-level metaclusters' scores by averaging their corresponding items.
The convergent validity of the final set of items was assessed by computing the correlations between the Mint scale and the other interoception questionnaires (MAIA, BPQ, IAS).

Next, we tested the predictive power of the Mint scale relative to other interoception questionnaires.
We will assessed the importance of each interoceptive dimension (from all the scales) as a unique predictor by fitting regression models (linear for continuous measures - e.g., depression score - and logistic for binary variables - e.g., presence *vs.* absence of mood disorders) to predict different outcomes, and compare the best between the 4 interoception scales (based on the R2).
We then evaluated the predictive performance of each scale as a whole by comparing regression models with all the dimensions entered as predictors (note that the IAS and BPQ only have one total score variable). We assessed the models based on their total explained variance (R2), as well as on the Bayesian Information Criterion (BIC), which penalizes for predictor number, thus offering a balance between model parsimony and predictive power. For the logistic models, we quantified the discriminative power by computing the Area Under the Curve (AUC) of the Receiver Operating Characteristic (ROC) curve, which assesses the model's discriminative power (the combination of sensitivity and specificity).

In order to evaluate the potential for a short version of our scale, we compared 4 variants of the Mint: the (full) *Mint* (including all lower-level dimensions), the *metaMint* (only including the metaclusters), the *miniMint* (including the metaclusters computed from a reduced number of items), and the *microMint* (including only the most representative dimension of each metacluster).
Moreover, we also included an alternative "interoception-focused" version of the MAIA (*iMAIA*) that only contains the 3 most interoceptive dimensions (Body Listening, Noticing, and Emotional Awareness)^[See correlation results to further justify this selection.].

@@

The application of EGA to the initial set of 42 items reproduced the expected 14 lower-level clusters of triplets and 3 higher-order metaclusters, with the exception of the *UrSe* (Urointestinal Sensitivity), for which one item moved to the *Olfa* (Olfactory Compensation) cluster.
In order to further balance and reduce the items, we removed the *UrSe* cluster, as well as *StaS* (State Specificity), which in comparison to the other items stood out as containing vague and overlapping items
<!-- ; e.g., "Being sexually aroused is a very different bodily feeling compared to other states (e.g., feeling anxious, relaxed, or after physical exercise)" -->
; *SexO* (Sexual Organs Sensitivity), which overlapped with the *SexS* (Sexual Arousal Sensitivity) cluster; and *CaNo* (Cardiorespiratory Noticing), which overlapped with the *CaCo* (Cardiorespiratory Confusion) cluster.

The final set of 33 items yielded a good fit (Generalized Total Entropy Fit Index = -77.23), with all items showing high cluster stability (> 90\%), with the exception of *Olfa*. The final structure included 11 lower-level clusters, grouped into 3 higher-order metaclusters: "Interoceptive Deficit" (containing *CaCo*, *UrIn*, *Derm*, *Sati*, *Olfa*), "Interoceptive Awareness" (containing *ExAc*, *RelA*, *SexS*), and "Visceroception" (containing *Card*, *Resp*, *Gast*). Item centrality (intepretable as loadings in traditional factor analysis frameworks) are shown in @fig-three. Additionally, we applied hierarchical clustering analysis which replicated this structure (see @fig-two), suggesting consistency across methods.
